---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## 1. Loading and preprocessing the data

We need to unzip data first:

```{r echo=TRUE}
unzip('./activity.zip')
df_activity <- read.csv('./activity.csv')
```

Let's take a look on the obtained data.

Let's look on the top items, items' datatypes and summary on the data:

```{r echo=TRUE}
head(df_activity)
```

```{r echo=TRUE}
str(df_activity)
```

```{r echo=TRUE}
summary(df_activity)
```

Column **`steps`** has NA values.

The datatypes seem to be allright.

## 2. What is mean total number of steps taken per day?

The first step is calculating the total number of steps taken per day. For this, we do the following steps:

Create data frame with the number of steps corresponding each date:

```{r echo=TRUE}
num_steps_per_day <- data.frame(aggregate(steps ~ date,
                                df_activity,
                                sum, 
                                # omit rows with missing values
                                na.rm = TRUE))
```

Check the obtained data frame:

```{r echo=TRUE}
head(num_steps_per_day)
```

Creating histogram on the total number of steps taken each day:

```{r echo=TRUE}

hist(num_steps_per_day$steps,
     main = "Total number of steps taken each day",
     xlab = "Steps",
     ylab = "Value",
     col = "deepskyblue3")
```

Calculating the mean of the total steps taken per day:

```{r echo=TRUE}
steps_per_day_mean <- mean(num_steps_per_day$steps)
steps_per_day_mean
```

Calculating the median of the total steps taken per day:

```{r echo=TRUE}
steps_per_day_median <- median(num_steps_per_day$steps)
steps_per_day_median
```

## 3. What is the average daily activity pattern?

## 4. Imputing missing values

## 5. Are there differences in activity patterns between weekdays and weekends?
