---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## 1. Loading and preprocessing the data

We need to unzip data first:

```{r echo=TRUE}
unzip('./activity.zip')
df_activity <- read.csv('./activity.csv')
```

Let's take a look on the obtained data.

Let's look on the top items, items' datatypes and summary on the data:

```{r echo=TRUE}
head(df_activity)
```

```{r echo=TRUE}
str(df_activity)
```

```{r echo=TRUE}
summary(df_activity)
```

Column **`steps`** has NA values.

The datatypes seem to be allright.

## 2. What is mean total number of steps taken per day?

The first step is calculating the total number of steps taken per day. For this, we do the following steps:

Create data frame with the number of steps corresponding each date:

```{r echo=TRUE}
num_steps_per_day <- data.frame(aggregate(steps ~ date,
                                df_activity,
                                sum, 
                                # omit rows with missing values
                                na.rm = T))
```

Check the obtained data frame:

```{r echo=TRUE}
head(num_steps_per_day)
```

Creating histogram on the total number of steps taken each day:

```{r echo=TRUE}
hist(num_steps_per_day$steps,
     main = "Total number of steps taken each day",
     xlab = "Steps",
     ylab = "Value",
     col = "deepskyblue3")
```

Calculating the mean of the total steps taken per day:

```{r echo=TRUE}
steps_per_day_mean <- mean(num_steps_per_day$steps)
steps_per_day_mean
```

Calculating the median of the total steps taken per day:

```{r echo=TRUE}
steps_per_day_median <- median(num_steps_per_day$steps)
steps_per_day_median
```

## 3. What is the average daily activity pattern?

We have to make a time series plot of 5-minute interval and then take average number of steps taken averaged across all days:

```{r echo=TRUE}
steps_per_5_min <- aggregate(steps ~ interval,
                             data = df_activity,
                             mean,
                             # omit rows with missing values
                             na.rm = T)
```

Plotting obtained time series:

```{r echo=TRUE}
plot(steps ~ interval,
     data = steps_per_5_min,
     type = "l",
     main = "Average daily activity pattern, steps per day",
     xlab = "Days",
     ylab = "Averaged number of steps")
```

Finding the 5-min interval, that contains the maximum number of steps:

```{r echo=TRUE}
steps_per_5_min_max <- steps_per_5_min[which.max(steps_per_5_min$steps),]$interval
steps_per_5_min_max
```

This 5 min interval (**835**) contains the maximum number of steps across the averaged intervals.

## 4. Imputing missing values

The first step is calculating the total number of missing values in the dataset.

```{r echo=TRUE}
na_number <- sum(is.na(df_activity))
na_number
```

The total number of missing values in the data set is **2304**.

The second step is filling the missing values. The best way is filling the gaps in the data with the *mean* values. In our case, we will fill the gaps with the *values of means per interval*. Let's create a function that returns a mean value per interval.

```{r echo=TRUE}
interval_mean <- function(interval){
        steps_per_5_min[steps_per_5_min$interval==interval,]$steps
}
```

Using the function above, let's create a gap-filling method itself. In a loop we check if the value is NA or not, in case if it's NA, we write instead of NA the value of the mean steps per interval using previously created function:

```{r echo=TRUE}
df_activity_no_na <- df_activity
for(i in 1:17568){
        if(is.na(df_activity_no_na[i,]$steps)){
                df_activity_no_na[i,]$steps <- interval_mean(df_activity_no_na[i,]$interval)
        }
}
```

To check if this worked properly and there are no NA values, let's count the NA values in the new data frame, like we did before:

```{r echo=TRUE}
na_number_2 <- sum(is.na(df_activity_no_na))
na_number_2
```

There are **no** NA values.

Finally, creating histogram with the no NA data frame in order to represent total number of steps, taken each day, like we did in the task 2.

Aggregating needed data:

```{r echo=TRUE}
num_steps_per_day_no_na <- data.frame(aggregate(steps ~ date,
                                      df_activity_no_na,
                                      sum))
```

Plotting histogram:

```{r echo=TRUE}
hist(num_steps_per_day_no_na$steps,
     main = "Total number of steps taken each day",
     xlab = "Steps",
     ylab = "Value",
     col = "deepskyblue2")
```

Calculating the mean of the total steps taken per day:

```{r echo=TRUE}
steps_per_day_mean <- mean(num_steps_per_day_no_na$steps)
steps_per_day_mean
```

Calculating the median of the total steps taken per day:

```{r echo=TRUE}
steps_per_day_median <- median(num_steps_per_day$steps)
steps_per_day_median
```

As we can deduce from the obtained results after filling NAs as a mean values of the steps on the interval, values of the mean and median never changed.

## 5. Are there differences in activity patterns between weekdays and weekends?

To have this task done, regards the fact the date is given as a `chr` datatype, we need to convert it to the right datatype first:

```{r echo=TRUE}
df_activity_no_na$date <- as.POSIXct(df_activity_no_na$date, 
                                     "%Y-%m-%d")
```

Checking if the data datype on the `date` columns is `POSIXct`. The column type is right:

```{r echo=TRUE}
str(df_activity_no_na)
```

To define if the selected date is weekday or weekend, we'll use `weekend()` function. The result of this is written to a new column `weekday`, the values Weekday or Weekend respectively to the day type the current date belongs to:

```{r echo=TRUE}
df_activity_no_na$weekday <- sapply(df_activity_no_na$date,
        function(x){
                        if(weekdays(x)=="Saturday"|weekdays(x)=="Sunday") {
        day <- "Weekend"
} else {
        day <- "Weekday"
}
                                    } )
```

Check, if the column created on the previous step is right. It's right and contains values Weekday or Weekend:

```{r echo=TRUE}
head(df_activity_no_na)
```

In order to make a plot, we have to aggregate values of the steps per interval, respecting to the factor of if the day is weekend or weekday (data from the column `weekday`) :

```{r echo=TRUE}
steps_by_day_of_week <- aggregate(steps ~ interval + weekday,
                                  df_activity_no_na,
                                  mean)
```

Creating a plot using `ggplot2` library in order to make the visualization more flexible.

From the obtained visualization we can deduce that average steps activity on the weekends is a little bit higher than on the weekdays.

```{r echo=TRUE}
library(ggplot2)
base_plot <- ggplot(steps_by_day_of_week, 
                    aes(x = interval, 
                        y = steps, 
                        color = weekday)) 
weekday_weekend_activity_plot <-  base_plot + 
                                  geom_line() + 
                                  ggtitle("Steps activity pattern for weekdays and weekends") + 
                                   xlab("Interval") + 
                                   ylab("AVG number of steps") +
        # creating panels
                                   facet_wrap(~weekday, ncol = 1, nrow = 2) +
                                   scale_color_discrete(name = "Day of the week")

print(weekday_weekend_activity_plot)
```
